You Draw It': Implementation of visually fitted trends with `r2d3`

### 2022-07-11 2022-sdss-you-draw-it-manuscript.qmd
updated template from SDSS email. Adjusted figures to display only one set.

keywords: 'graphics, user interaction, regression'
format: 
  pdf:
    include-in-header:
      - edits.tex
      - refs.tex
    keep-tex: true
    # template: template-blinded.tex
    # template: template.tex
---

@@

library(readxl)
library(lubridate)
```

@@

eyefitting_example_sim <- read.csv("data/eyefitting-simdata-example.csv")
```

@@

We conducted a study in order to validate 'You Draw It' as a method for graphical testing, comparing results to the less technological method utilized in @mosteller1981eye.
The original study asked 153 graduate students and postdoctoral researchers in an Introductory Biostatistics course to fit lines by eye to a set of four points (S, F, V, N), each set with differing slope and variance properties, using an 8.5 x 11 inch transparency with a straight line etched across the middle.
Researchers conducted the study with a latin square experimental design to test for a practice effect on the task.
Results indicated participants tended to fit the slope of the first principal component (minimizes orthogonal residuals) as opposed to the slope from the ordinary least squares regression (minimizes vertical residuals) [@fig-pca-plot] and found no effect of order.

```{r ols-vs-pca-example, message = F, warning = F}
#| fig-cap: "Comparison between an OLS regression equation which minimizes the vertical distance of points from the line and a regression equation with a slope calculated by the first principal component which minimizes the smallest distance of points from the line."
#| out-width: \columnwidth
#| label: fig-pca-plot
#| fig-res: 300
#| fig-width: 6
#| fig-height: 4
#| fig-env: figure

@@

In our study, we replicated @mosteller1981eye using the 'You Draw It' method and simulated data with parameter coefficients selected to reflect those from the four data sets in the original study. Data were simulated based on a linear model with additive errors:
\begin{align}
y_i & = \beta_0 + \beta_1 x_i + e_i \\
\text{with } e_i & \sim N(0, \sigma^2). \nonumber
\end{align} 
For each participant, a unique data set was randomly and independently generated from the underlying parameters at the beginning of the study and mapped to a scatter-plot graphic.
@fig-eyefitting-simplot illustrates an example of simulated data for all four parameter choices intended to reflect the trends in the original study.
When participants started the study, they were first asked to complete two practice plots - accompanied by instructions and a .gif demonstrating the task - in order to train them in the skills associated with executing the task.
The four 'You Draw It' task plots associated with the selected parameters followed the practice plots in a random order for each individual.

```{r eyefitting-simplot, warning = F, message = F}
#| fig-cap: "Example of simulated data points displayed in a scatter-plot illustrating the trends associated with the four selected parameter choices."
#| out-width: \columnwidth
#| label: fig-eyefitting-simplot
#| fig-res: 300
#| fig-width: 9
#| fig-height: 9
#| fig-env: figure

eyefitting_example_simplot <- eyefitting_example_sim %>%
  filter(data == "point_data") %>%
  filter(dataset %in% c("F", "N", "S") | (x < 16 & x > 4)) %>%
  mutate(dataset = factor(dataset, levels = c("S", "F", "V", "N"))) %>%
  dplyr::rename(`Parameter Choice` = dataset) %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(size = 1) +
  facet_wrap(~`Parameter Choice`, ncol = 2) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
  legend.position = "none",
  plot.title   = element_text(size = 12, hjust = 0),
  axis.text    = element_text(size = 12),
  axis.title   = element_text(size = 12),
  legend.title = element_text(size = 12),
  legend.text  = element_text(size = 12),
  # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
  # strip.background = element_rect(size = 0.5),
  legend.key.size = unit(1, "line")
) +
  scale_y_continuous(breaks = seq(-10, 20, 5))
eyefitting_example_simplot
```

Participants were recruited via Prolific in March 2022; while this study does utilize a convenience sample, as this is primarily a perceptual task, previous results have found few differences between expert and non-expert participants in this context [CITE SPATIAL PAPER].
The study was conducted and distributed via a Shiny application and participants completed the tasks on their own computer, with a computer mouse, and in an environment of their choosing.
The data from this study were collected to validate this method of graphical testing, with the hopes of providing a new tool to assess graphical perception interactively.

@@

As a part of the validation study, we collected data for each participant drawn trend line, allowing us to conduct formal analyses which compared the intuitive visually fit lines to statistical regression results - the original study lacked these formal comparisons, providing only summaries of averages, variances, and actual (least squares) values for the slope and intercept of each data set.
The feedback data, stored in a SQLite database, contained the simulated data points} - $(x_{ijk}, y_{ijk})$ -, the predicted values from the statistical regression models - $\hat y_{ijk,OLS}$, and $\hat y_{ijk,PCA}$ -, and the predicted values from the user drawn line - $y_{ijk,drawn}$ for parameter choice $i = 1,2,3,4$, $j = 1,...N_\text{participant}$, and $x_{ijk}$ value $k = 1, ...,4 x_{max} + 1$.
@fig-eyefitting-trial-plot displays an example of all three fitted trend lines for parameter choice F.

eyefitting_example <- eyefitting_model_data %>%
  # filter(participant_id == "65c10d498eae365e108efcd3dcb75287", parm_id == "N") %>%
  filter(participant_id == "60b16b9bd5a122c1457d31055df51a45", `Parameter Choice` == "F") %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = yols, color = "OLS", linetype = "OLS")) +
  geom_line(aes(y = ypca, color = "PCA", linetype = "PCA")) +
  geom_line(aes(y = ydrawn, color = "Drawn", linetype = "Drawn")) +
  geom_point(data = eyefitting_simulated_data %>%
               filter(dataset == "point_data", participant_id == "60b16b9bd5a122c1457d31055df51a45", `Parameter Choice` == "F"),
             aes(x = x, y = y)) +
  facet_wrap(~`Parameter Choice`, labeller = labeller(`Parameter Choice` = label_both)) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "bottom",
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 8, margin = margin(0.1,0,0.1,0, "cm")),
        # strip.background = element_rect(size = 0.8),
        legend.key.size = unit(1, "line")
        ) +
  scale_x_continuous(limits = c(0,20)) +
  scale_color_manual("", values = c("black", "steelblue", "orange")) +
  scale_linetype_manual("", values = c("dashed", "solid", "solid")) +
  scale_y_continuous("y")

eyefitting_example
```

@@

@@

# Plot Predictions
eyefitting_lmer_plot <- eyefitting_lmer_preds %>%
  filter((`Parameter Choice` %in% c("F", "N", "S") | (x <= 16 & x >= 4))) %>%
  mutate(`Parameter Choice` = factor(`Parameter Choice`, levels = c("S", "F", "V", "N"))) %>%
  ggplot(aes(x = x)) +
  geom_line(data = eyefitting_model_data %>% 
              filter(prolific_id %in% participant_sample), 
            aes(x = x, y = residual_ols_loess, group = plot_id, color = "OLS"), alpha = 0.1) +
  geom_line(data = eyefitting_model_data %>% 
              filter(prolific_id %in% participant_sample), 
            aes(x = x, y = residual_pca_loess, group = plot_id, color = "PCA"), alpha = 0.1) +
  geom_ribbon(aes(ymin = asymp.LCL.ols, ymax = asymp.UCL.ols, fill = "OLS"), color = NA, alpha = 0.4) +
  geom_line(aes(y = emmean.ols, color = "OLS")) +
  geom_ribbon(aes(ymin = asymp.LCL.pca, ymax = asymp.UCL.pca, fill = "PCA"), color = NA, alpha = 0.4) +
  geom_line(aes(y = emmean.pca, color = "PCA")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~`Parameter Choice`, labeller = labeller(`Parameter Choice` = label_both), scales = "free") +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "right",
        plot.title   = element_text(size = 12, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
        # strip.background = element_rect(size = 0.5),
        legend.key.size = unit(1, "line")
        ) +
  scale_y_continuous("Residual", limits = c(-5,5), breaks = seq(-5,5,2.5)) +
  scale_color_manual("Individual participant \nresiduals", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) +
  scale_fill_manual("LMER fitted trend", values = c("steelblue", "orange"), labels = c("OLS", "PCA"))  +
  guides(color = guide_legend(override.aes = list(alpha = 1)),
         fill = guide_legend(override.aes = list(alpha = 1)))

eyefitting_lmer_plot
```

@@

eyefitting_gamm_plot <- eyefitting_gamm_preds %>%
  filter((`Parameter Choice` %in% c("F", "N", "S") | (x <= 16 & x >= 4))) %>%
  mutate(`Parameter Choice` = factor(`Parameter Choice`, levels = c("S", "F", "V", "N"))) %>%
  ggplot(aes(x = x)) +
  geom_line(data = eyefitting_model_data %>% 
              filter(prolific_id %in% participant_sample), 
            aes(x = x, y = residual_ols_loess, group = plot_id, color = "OLS"), alpha = 0.1) +
  geom_line(data = eyefitting_model_data %>% 
              filter(prolific_id %in% participant_sample), 
            aes(x = x, y = residual_pca_loess, group = plot_id, color = "PCA"), alpha = 0.1) +
  geom_ribbon(aes(ymin = ols.lower, ymax = ols.upper, fill = "OLS"), color = NA, alpha = 0.4) +
  geom_line(aes(y = ols.pred, color = "OLS")) +
  geom_ribbon(aes(ymin = pca.lower, ymax = pca.upper, fill = "PCA"), color = NA, alpha = 0.4) +
  geom_line(aes(y = pca.pred, color = "PCA")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~`Parameter Choice`, labeller = labeller(`Parameter Choice` = label_both), scales = "free") +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "right",
        plot.title   = element_text(size = 12, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
        # strip.background = element_rect(size = 0.5),
        legend.key.size = unit(1, "line")
        ) +
  scale_y_continuous("Residual", limits = c(-5,5), breaks = seq(-5,5,2.5)) +
  scale_color_manual("Individual participant \nresiduals", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) +
  scale_fill_manual("GAMM fitted trend", values = c("steelblue", "orange"), labels = c("OLS", "PCA"))   +
  guides(color = guide_legend(override.aes = list(alpha = 1)),
         fill = guide_legend(override.aes = list(alpha = 1)))

eyefitting_gamm_plot
```

@fig-eyefitting-lmer-residualplots and @fig-eyefitting-gamm-residualplots show a visual analysis of the estimated trends of residuals (vertical deviation of participant drawn points from both the OLS and PCA fitted points) as modeled by a LMM and GAMM respectively. 
A random sample of 75 participants was selected to display individual participant residuals behind the overall residual trend.
Examining the plots, the estimated trends of PCA residuals (orange) appear to align more parallel and closer to the $y=0$ horizontal (dashed) line than the OLS residuals (blue). 
In particular, this trend is more prominent in parameter choices with large variances (F and N).
Results from our study were consistent with those found in the original study; when shown points following a linear trend, participants tended to fit the slope of the first principal component Our study established 'You Draw It' as a method for graphical testing and reinforced the differences between intuitive visual model fitting and statistical model fitting, providing information about human perception as it relates to the use of statistical graphics.

@@

After validating the 'You Draw It' method, we conducted a study utilizing the new method to evaluate participants ability to make forecasts for exponentially increasing data on log and linear scales. Along with analyzing the feedback data with the GAMM method for flexibility due to nonlinear data, we used spaghetti plots to conduct a visual analysis of participant forecasts compared to the nonlinear least squares statistical model and make comparisons between two chart design features [@fig-exponential-yloess-spaghetti-plot-2-1]. This study demonstrated the advantage of using the 'You Draw It' method paired with GAMM and visual analysis methods for measuring perception and visual patterns of nonlinear data.

![Spaghetti plot of results from a study which asked participants to forcast trends of exponentially increasing data. Participants drawn lines on the linear scale are shown in blue and the log scale are shown in orange. Variability in the statistically fitted regression lines occured due to a unique data set being simulated for each individual; the gray band shows the range fitted values from the statistically fitted regression lines.](images/exponential-yloess-spaghetti-plot-2-1.png){#fig-exponential-yloess-spaghetti-plot-2-1}

>>>>

keywords: 'graphics, user interaction, regression'
format: 
  pdf:
    include-in-header:
      - edits.tex
      # - refs.tex
    keep-tex: true
    # template: template.tex
documentclass: jds
---

@@

library(readxl)
library(lubridate)
library(patchwork)
```

@@

eyefitting_example_sim <- read.csv("data/eyefitting-simdata-example.csv")

youdrawit_model_data      <- read_csv("data/youdrawit-model-data.csv")  %>%
  mutate(`Points Truncated` = ifelse(points_truncated == 10, "50%", "75%"),
         `Growth Rate` = ifelse(beta == "beta0.1", "Low", "High")) %>%
  mutate(`Growth Rate` = factor(`Growth Rate`, levels = c("Low", "High"))) %>%
  filter(`Points Truncated` == "50%", `Growth Rate` == "High")

youdrawit_simulated_band <- youdrawit_model_data %>%
  group_by(`Growth Rate`, `Points Truncated`, x) %>%
  summarize(min_ynls = min(ynls),
            max_ynls = max(ynls))
```

```{r eyefitting-simplot, eval = F, warning = F, message = F}
#| fig-cap: "Example of simulated data points displayed in a scatter-plot illustrating the trends associated with the four selected parameter choices."
#| out-width: \columnwidth
#| label: fig-eyefitting-simplot
#| fig-res: 300
#| fig-width: 9
#| fig-height: 9
#| fig-env: figure

eyefitting_example_simplot <- eyefitting_example_sim %>%
  filter(data == "point_data") %>%
  filter(dataset %in% c("F", "N", "S") | (x < 16 & x > 4)) %>%
  mutate(dataset = factor(dataset, levels = c("S", "F", "V", "N"))) %>%
  filter(`Parameter Choice` == "F") %>%
  dplyr::rename(`Parameter Choice` = dataset) %>%
  ggplot(aes(x = x, y = y)) +
  geom_point(size = 1) +
  facet_wrap(~`Parameter Choice`, ncol = 2) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
  legend.position = "none",
  plot.title   = element_text(size = 12, hjust = 0),
  axis.text    = element_text(size = 12),
  axis.title   = element_text(size = 12),
  legend.title = element_text(size = 12),
  legend.text  = element_text(size = 12),
  # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
  # strip.background = element_rect(size = 0.5),
  legend.key.size = unit(1, "line")
) +
  scale_y_continuous(breaks = seq(-10, 20, 5))
# eyefitting_example_simplot
```

eyefitting_example <- eyefitting_model_data %>%
  # filter(participant_id == "65c10d498eae365e108efcd3dcb75287", parm_id == "N") %>%
  filter(participant_id == "60b16b9bd5a122c1457d31055df51a45", `Parameter Choice` == "F") %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = yols, color = "OLS", linetype = "OLS")) +
  geom_line(aes(y = ypca, color = "PCA", linetype = "PCA")) +
  geom_line(aes(y = ydrawn, color = "Drawn", linetype = "Drawn")) +
  geom_point(data = eyefitting_simulated_data %>%
               filter(dataset == "point_data", participant_id == "60b16b9bd5a122c1457d31055df51a45", `Parameter Choice` == "F"),
             aes(x = x, y = y)) +
  facet_wrap(~`Parameter Choice`, labeller = labeller(`Parameter Choice` = label_both)) +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "bottom",
        plot.title   = element_text(size = 14, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 8, margin = margin(0.1,0,0.1,0, "cm")),
        # strip.background = element_rect(size = 0.8),
        legend.key.size = unit(1, "line")
        ) +
  scale_x_continuous(limits = c(0,20)) +
  scale_color_manual("", values = c("black", "steelblue", "orange")) +
  scale_linetype_manual("", values = c("dashed", "solid", "solid")) +
  scale_y_continuous("y")

# eyefitting_example
```

# Plot Predictions
eyefitting_lmer_plot <- eyefitting_lmer_preds %>%
  filter((`Parameter Choice` %in% c("F", "N", "S") | (x <= 16 & x >= 4))) %>%
  mutate(`Parameter Choice` = factor(`Parameter Choice`, levels = c("S", "F", "V", "N"))) %>%
  filter(`Parameter Choice` == "F") %>%
  ggplot(aes(x = x)) +
  geom_line(data = eyefitting_model_data %>% 
              filter(prolific_id %in% participant_sample)  %>% filter(`Parameter Choice` == "F"), 
            aes(x = x, y = residual_ols_loess, group = plot_id, color = "OLS"), alpha = 0.1) +
  geom_line(data = eyefitting_model_data %>% 
              filter(prolific_id %in% participant_sample) %>% filter(`Parameter Choice` == "F"), 
            aes(x = x, y = residual_pca_loess, group = plot_id, color = "PCA"), alpha = 0.1) +
  geom_ribbon(aes(ymin = asymp.LCL.ols, ymax = asymp.UCL.ols, fill = "OLS"), color = NA, alpha = 0.4) +
  geom_line(aes(y = emmean.ols, color = "OLS")) +
  geom_ribbon(aes(ymin = asymp.LCL.pca, ymax = asymp.UCL.pca, fill = "PCA"), color = NA, alpha = 0.4) +
  geom_line(aes(y = emmean.pca, color = "PCA")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~`Parameter Choice`, labeller = labeller(`Parameter Choice` = label_both), scales = "free") +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "bottom",
        plot.title   = element_text(size = 14, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
        # strip.background = element_rect(size = 0.5),
        legend.key.size = unit(1, "line"), legend.box="vertical", legend.margin=margin()
        ) +
  scale_y_continuous("Residual", limits = c(-5,5), breaks = seq(-5,5,2.5)) +
  scale_color_manual("Individual participant \nresiduals", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) +
  scale_fill_manual("LMER fitted trend", values = c("steelblue", "orange"), labels = c("OLS", "PCA"))  +
  guides(color = guide_legend(override.aes = list(alpha = 1)),
         fill = guide_legend(override.aes = list(alpha = 1)))

# eyefitting_lmer_plot
```

eyefitting_gamm_plot <- eyefitting_gamm_preds %>%
  filter((`Parameter Choice` %in% c("F", "N", "S") | (x <= 16 & x >= 4))) %>%
  mutate(`Parameter Choice` = factor(`Parameter Choice`, levels = c("S", "F", "V", "N"))) %>%
  filter(`Parameter Choice` == "F") %>%
  ggplot(aes(x = x)) +
  geom_line(data = eyefitting_model_data %>% 
              filter(prolific_id %in% participant_sample)  %>% filter(`Parameter Choice` == "F"), 
            aes(x = x, y = residual_ols_loess, group = plot_id, color = "OLS"), alpha = 0.1) +
  geom_line(data = eyefitting_model_data %>% 
              filter(prolific_id %in% participant_sample)  %>% filter(`Parameter Choice` == "F"), 
            aes(x = x, y = residual_pca_loess, group = plot_id, color = "PCA"), alpha = 0.1) +
  geom_ribbon(aes(ymin = ols.lower, ymax = ols.upper, fill = "OLS"), color = NA, alpha = 0.4) +
  geom_line(aes(y = ols.pred, color = "OLS")) +
  geom_ribbon(aes(ymin = pca.lower, ymax = pca.upper, fill = "PCA"), color = NA, alpha = 0.4) +
  geom_line(aes(y = pca.pred, color = "PCA")) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~`Parameter Choice`, labeller = labeller(`Parameter Choice` = label_both), scales = "free") +
  theme_bw(base_size = 14) +
  theme(aspect.ratio = 1,
        legend.position = "bottom",
        plot.title   = element_text(size = 14, hjust = 0),
        axis.text    = element_text(size = 12),
        axis.title   = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text  = element_text(size = 12),
        # strip.text = element_text(size = 5, margin = margin(0.05,0,0.05,0, "cm")),
        # strip.background = element_rect(size = 0.5),
        legend.key.size = unit(1, "line"), legend.box="vertical", legend.margin=margin()
        ) +
  scale_y_continuous("Residual", limits = c(-5,5), breaks = seq(-5,5,2.5)) +
  scale_color_manual("Individual participant \nresiduals", values = c("steelblue", "orange"), labels = c("OLS", "PCA")) +
  scale_fill_manual("GAMM fitted trend", values = c("steelblue", "orange"), labels = c("OLS", "PCA"))   +
  guides(color = guide_legend(override.aes = list(alpha = 1)),
         fill = guide_legend(override.aes = list(alpha = 1)))

# eyefitting_gamm_plot
```

@@

We conducted a study in order to validate 'You Draw It' as a method for graphical testing, comparing results to the less technological method utilized in @mosteller1981eye.
The original study asked 153 graduate students and postdoctoral researchers in an Introductory Biostatistics course to fit lines by eye to a set of four points (S, F, V, N), each set with differing slope and variance properties, using an 8.5 x 11 inch transparency with a straight line etched across the middle.
Researchers conducted the study with a latin square experimental design to test for a practice effect on the task.
Results indicated participants tended to fit the slope of the first principal component (minimizes Euclidean distance) as opposed to the slope from the ordinary least squares regression (minimizes vertical distance) [@fig-pca-plot] and found no effect of order.

```{r ols-vs-pca-example, message = F, warning = F}
#| fig-cap: "Comparison between an OLS regression equation which minimizes the vertical distance of points from the line and a regression equation with a slope calculated by the first principal component which minimizes the smallest distance of points from the line."
#| out-width: "70%"
#| label: fig-pca-plot
#| fig-res: 300
#| fig-width: 6
#| fig-height: 4
#| fig-env: figure

@@

In our study, we replicated @mosteller1981eye using the 'You Draw It' method and simulated data with parameter coefficients selected to reflect those from the four data sets in the original study. Data were simulated based on a linear model with additive errors:
\begin{align}
y_i & = \beta_0 + \beta_1 x_i + e_i \\
\text{with } e_i & \sim N(0, \sigma^2). \nonumber
\end{align} 
For each participant, four unique data sets were randomly and independently generated from the underlying parameters at the beginning of the study and mapped to a scatter-plot graphic.
<!-- @fig-eyefitting-simplot illustrates an example of simulated data for all four parameter choices intended to reflect the trends in the original study. -->
When participants started the study, they were first asked to complete two practice plots - accompanied by instructions and a .gif demonstrating the task - in order to train them in the skills associated with executing the task.
The four 'You Draw It' task plots associated with the selected parameters followed the practice plots in a random order for each individual.

Participants were recruited via Prolific in March 2022; while this study does utilize a convenience sample, as this is primarily a perceptual task, previous results have found few differences between expert and non-expert participants in this context \ear{(CITE SPATIAL PAPER)}.
The study was conducted and distributed via a Shiny application and participants completed the tasks on their own computer, with a computer mouse, and in an environment of their choosing.
The data from this study were collected to validate this method of graphical testing, with the hopes of providing a new tool to assess graphical perception interactively.

@@

As a part of the validation study, we collected data for each participant drawn trend line, allowing us to conduct formal analyses which compared the intuitive visually fit lines to statistical regression results - the original study lacked these formal comparisons, providing only summaries of averages, variances, and actual (least squares) values for the slope and intercept of each data set.
The feedback data, stored in a SQLite database, contained the simulated data points} - $(x_{ijk}, y_{ijk})$ -, the predicted values from the statistical regression models - $\hat y_{ijk,OLS}$, and $\hat y_{ijk,PCA}$ -, and the predicted values from the user drawn line - $y_{ijk,drawn}$ for parameter choice $i = 1,2,3,4$, $j = 1,...N_\text{participant}$, and $x_{ijk}$ value $k = 1, ...,4 x_{max} + 1$.
@fig-eyefitting-plots displays an example of all three fitted trend lines for parameter choice F.

@@

```{r eyefitting-plots, warning = F, message = F}
#| fig-cap: "Visual analysis results for parameter choice set F. \\linebreak \\textbf{A.} Example of validation feedback data where three trend lines show the the OLS fitted, PCA fitted, and participant drawn values overlaid on the simulated data points. \\linebreak \\textbf{B \\& C.} Estimated trends of residuals (vertical deviation of participant drawn points from both the OLS (blue) and PCA (orange) fitted points) as fit by the (B) linear mixed model and (C) generalized additive mixed model. A random sample of 75 participants was selected to display the individual participant residuals behind the overall trend."
#| out-width: \columnwidth
#| label: fig-eyefitting-plots
#| fig-res: 300
#| fig-width: 12
#| fig-height: 6
#| fig-env: figure

(eyefitting_example + ggtitle("A. Example")) + 
  (eyefitting_lmer_plot + ggtitle("B. LMER Results")) + 
  (eyefitting_gamm_plot + ggtitle("C. GAMM Results")) +
  plot_layout(ncol =  3)
```

@@

@@

While the statistical models were fit to all four parameter choices, @fig-eyefitting-plots displays the estimated trends of residuals (vertical deviation of participant drawn points from both the OLS and PCA fitted points) as modeled by a LMM and GAMM respectively for parameter choice set F. 
A random sample of 75 participants was selected to display individual participant residuals behind the overall residual trend.
Examining the plots, the estimated trends of PCA residuals (orange) appear to align more parallel and closer to the $y=0$ horizontal (dashed) line than the OLS residuals (blue). 
In particular, this trend was more prominent in parameter choices with large variances.
Results from our study were consistent with those found in the original study; when shown points following a linear trend, participants tended to fit the slope of the first principal component Our study established 'You Draw It' as a method for graphical testing and reinforced the differences between intuitive visual model fitting and statistical model fitting, providing information about human perception as it relates to the use of statistical graphics.

@@

After validating the 'You Draw It' method, we conducted a study utilizing the new method to evaluate participants ability to make forecasts for exponentially increasing data on log and linear scales. Along with analyzing the feedback data with the GAMM method for flexibility due to nonlinear data, we used spaghetti plots to conduct a visual analysis of participant forecasts compared to the nonlinear least squares statistical model and make comparisons between two chart design features [@fig-exponential-spaghetti-plot]. This study demonstrated the advantage of using the 'You Draw It' method paired with GAMM and visual analysis methods for measuring perception and visual patterns of nonlinear data.

```{r eyefitting-plots, warning = F, message = F}
#| fig-cap: "Spaghetti plot of results from a study which asked participants to forecast trends of exponentially increasing data. Participants drawn lines on the linear scale are shown in blue and the log scale are shown in orange. Variability in the statistically fitted regression lines occurred due to a unique data set being simulated for each individual; the gray band shows the range fitted values from the statistically fitted regression lines."
#| out-width: \columnwidth
#| label: fig-exponential-spaghetti-plot
#| fig-res: 300
#| fig-width: 6
#| fig-height: 4
#| fig-env: figure

spaghetti_plot <- youdrawit_model_data %>%
  ggplot(aes(x = x)) +
  geom_line(aes(y = yloess, group = plot_id, color = scale), alpha = 0.2) +
  geom_ribbon(data = youdrawit_simulated_band, aes(ymin = min_ynls, ymax = max_ynls, fill = "Fitted NLS", group = NA), color = NA, alpha = 0.35) +
  # facet_grid(`Growth Rate` ~ `Points Truncated`, scales = "free", labeller = labeller(`Growth Rate` = label_both, `Points Truncated` = label_both)) +
  theme_bw() +
  theme(aspect.ratio = 1,
        legend.position = "right"
        ) +
  scale_color_manual("", values = c("steelblue", "orange"), labels = c("Visual fit, linear scale \n (drawn, loess)", "Visual fit, log scale \n (drawn, loess)")) +
  scale_fill_manual("", values = c("black"), labels = c("Range of fitted values \n from statistically fitted regression \n lines (NLS)")) +
  scale_x_continuous(limits = c(10, 20)) +
  scale_y_continuous("y", limits = c(0, 300), breaks = seq(0,300,50)) +
  guides(color = guide_legend(override.aes = list(alpha = 1)),
         fill = guide_legend(override.aes = list(alpha = 0.3)))
spaghetti_plot
```
